---
title: "Untitled"
author: "Emorie Beck"
date: "9/14/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Workspace
##Packages

```{r, message = F, results = 'hide', warning = F}
library(plyr)
library(tidyverse)
library(psych)
library(googlesheets)
library(lme4)
library(gridExtra)
library(stargazer)
library(stringr)
```

##Working Directory
```{r}
data_path <- "https://github.com/longitudinal-data/1-descriptives-and-graphs-emoriebeck/assigment2/raw/master"
```

1.Run linear models on all of your subjects (a basic regression). What is the average intercept, the average slope?

```{r}
load(url(paste(data_path, "data.RData", sep = "/")))
```

Now run a mlm/lmer model with only a random intercept. What is the ICC? What does residual variance look like compared to linear model? Create a graph to show this effect.

Introduce a fixed slope term. What is the difference in terms of the fixed effects estimates between this estimate and the previous? Of the residual standard error? Create a graph to show both fixed effects estimates and the CIs around them.

Run an additional model with a random slope. How does this change compare to the previous model? Should you keep the random slope or not?

Interpret the correlation between the slope and the intercept.

Create a density plot of the random effects from your final model.

Create a catepilar plot of the random effects. Is there any person that seems odd in terms of a large standard errors around intercept and slope estimates?

Create a plot of the trajectory, along with a spaghetti plot of each person’s individual slope. Set the alpha level (transparency) on the individual slopes to make them easier to see.

Create a plot of the trajectory, along with a spagehtti plot of each person’s individual slope. Set the alpha level (transperancy) on the individual slopes to make them easier to see.