---
title: "Assignment 3"
author: "Emorie D Beck"
date: "10/11/2017"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Workspace
##Packages
```{r, results = 'hide', message = F, warning = F}
library(psych)
library(sjPlot)
library(broom)
library(lme4)
library(MuMIn)
library(merTools)
library(reghelper)
library(stargazer)
library(lsmeans)
library(multcompView)
library(plyr)
library(tidyverse)
```


# Background
## Basic Syntex
From last week:
\begin{itemize}
  \item \textbf{Level 1:} $Y_{ij} = \beta_{0j} + \varepsilon{ij}$
  \item \textbf{Level 2:} $\beta_{0j} = \gamma_{00} + U_{0j}$
\end{itemize}

## Sample Data
The National Longitudinal Study of Youths 1979 Child and Young Adult Sample (NLSYCYA) is a longitudinal study conducted by the National Bureau of Labor Statistics. The sample includes the children of the original 1979 sample. Here, we are going to use a subset of the more than 11,000 variables available that include the following.

Item Name   | Description                   | Time-Varying?
----------- | ----------------------------- | -------------
PROC_CID    | Participant ID                | No  
Dem_DOB     | Year of Date of Birth         | No
groups      | Jail, Community Service, None | No
DemPWeight  | Weight Percentile at age 10   | No
age         | Age of participant            | Yes
Year        | Year of Survey                | Yes
age0        | Age of participant (centered) | Yes
SensSeek    | Sensation-Seeking Composite   | Yes
CESD        | CESD Depression Composite     | Yes

```{r, eval = T, tidy = T}
data_path <- "https://github.com/longitudinal-data/1-descriptives-and-graphs-emoriebeck/raw/master/Conditional_Models"
load(url(paste(data_path, "sample.RData", sep = "/")))

head(sample_dat)
```

# Time Invariant Predictors
## Time Invariant Predictors: Continuous

The basic equation, specifying a random intercept and slope:    
\begin{itemize}
  \item \textbf{Level 1:} $Y_{ij} = \beta_{0j} + \beta_{1j}*time_{1j} + \varepsilon{ij}$
  \item \textbf{Level 2:} 
    \begin{itemize} 
      \item $\beta_{0j} = \gamma_{00} + \gamma_{01}*X_{2j} + U_{0j}$
      \item $\beta_{1j} = \gamma_{10} + \gamma_{11}*X_{2j} + U_{1j}$
    \end{itemize}
\end{itemize}

But we need to break this down to see that adding additional predictors results in interaction terms:  

$Y_{ij} = \gamma_{00} + \gamma_{01}*X_{2j} + U_{0j} + (\gamma_{10} + \gamma_{11}*X_{2j} + U_{1j})*X_{1j} + \varepsilon{ij}$
$Y_{ij} = \gamma_{00} + \gamma_{01}*X_{2j} + \gamma_{10}*X_{1j} + \textcolor{red}{\gamma_{11}*X_{2j}*X_{1j}} + U_{0j} + U_{1j}*X_{1j} + \varepsilon{ij}$

We can also fit this with intercepts depending on weight, but without the change (slope) dependent on weight:  
$Y_{ij} = \gamma_{00} + \gamma_{01}*X_{2j} + U_{0j} + (\gamma_{10} + U_{1j})*X_{1j} + \varepsilon{ij}$
$Y_{ij} = \gamma_{00} + \gamma_{01}*X_{2j} + \gamma_{10}*X_{1j} + U_{0j} + U_{1j}*X_{1j} + \varepsilon{ij}$

### Continuous Example - Weight for Age Percentile  

\small
```{r}
describe(sample_dat$DemPweight)
```

\small
```{r,results='hide'}
# time invariant covariate with random intecept (with weight as covariate) 
# and slope (without weight as a covariate)
mod1a <- lmer(SensSeek ~ age0 + DemPweight + (age0|PROC_CID), 
              data = sample_dat)

summary(mod1a)

# time invariant predictor with random slope and intercept
mod1b <- lmer(SensSeek ~ age0 + DemPweight + age0*DemPweight + 
                (age0|PROC_CID), data = sample_dat)

summary(mod1b)
```

```{r}
fixed.frame <- sample_dat %>%
  summarise(mean = mean(DemPweight, na.rm = T), 
            sd = sd(DemPweight, na.rm = T))

fixed.frame <- 
  data.frame(
    expand.grid(
      # here, you add values for your time variable and predictors
      age0 = seq(0,10,2), 
      DemPweight = c(fixed.frame$mean-fixed.frame$sd,
                     fixed.frame$mean,
                     fixed.frame$mean+fixed.frame$sd))) %>%
  mutate(pred = predict(mod1b, newdata = ., re.form = NA))


fixed.frame %>%
  mutate(Weight = factor(DemPweight, levels = unique(DemPweight), labels = c("-1SD", "0SD", "1SD")),
         age = age0 + 14) %>%
  ggplot(aes(x = age, y = pred, color = Weight)) +
    geom_line(size = 2) + 
    labs(x = "Age", y = "Sensation Seeking Composite",
         title = "Continuous Invariant Conditional Growth Models") +
    theme_classic() +
    theme(axis.text = element_text(face = "bold", size = rel(1.2)),
          axis.title = element_text(face = "bold", size = rel(1.2)),
          legend.title = element_text(face = "bold", size = rel(1.2)),
          plot.title = element_text(face = "bold", size = rel(1.2), hjust = .5))
```

\normalsize



##Time Invariant Predictors: Categorical 
### Categorical Example - 2 level group

Let's start with the basic syntax:  

\begin{itemize}
  \item \textbf{Level 1:} $Y_{ij} = \beta_{0j} + \beta_{1j}*time_{1j} + \varepsilon{ij}$
  \item \textbf{Level 2:} 
    \begin{itemize} 
      \item $\beta_{0j} = \gamma_{00} + \gamma_{01}*X_{2j} + U_{0j}$
      \item $\beta_{1j} = \gamma_{10} + \gamma_{11}*X_{2j} + U_{1j}$
    \end{itemize}
\end{itemize}

Now let's swap that out for a 2 group sample from the present data:  

\begin{itemize}
  \item \textbf{Level 1:} $Y_{ij} = \beta_{0j} + \beta_{1j}*age0_{ij} + \varepsilon{ij}$
  \item \textbf{Level 2:} 
    \begin{itemize} 
      \item $\beta_{0j} = \gamma_{00} + \gamma_{01}*groupsNone + U_{0j}$
      \item $\beta_{1j} = \gamma_{10} + \gamma_{11}*groupsNone + U_{1j}$
    \end{itemize}
\end{itemize}

Variable  | D1  
--------- | ---
Jail      | 0 
None      | 1


And model it:  
\small
```{r}
mod2ga <- lmer(SensSeek ~ age0 + groups + (age0|PROC_CID), 
              data = sample_dat %>% filter(groups != "CommServ"))
mod2g <- lmer(SensSeek ~ age0 + groups + age0*groups + (age0|PROC_CID), 
              data = sample_dat %>% filter(groups != "CommServ"))
summary(mod2g)
```

\normalsize

```{r}
fixed.frame <- 
  data.frame(expand.grid(age0 = seq(0,10,2), 
             groups = c("None","Jail"))) %>%
  mutate(pred = predict(mod2g, newdata = ., re.form = NA))

fixed.frame %>%
  mutate(age = age0 + 14) %>%
  ggplot(aes(x = age, y = pred, color = groups)) +
    geom_line(size = 2) + 
    lims(y = c(0,4)) +
    labs(x = "Age", y = "Sensation Seeking Composite",
         title = "2 Group Time Invariant Conditional Growth Models") +
    theme_classic() +
    theme(axis.text = element_text(face = "bold", size = rel(1.2)),
          axis.title = element_text(face = "bold", size = rel(1.2)),
          legend.title = element_text(face = "bold", size = rel(1.2)),
          plot.title = element_text(face = "bold", size = rel(1.2), hjust = .5))

```

##Both Continuous and Categorical
```{r}
mod3 <- lmer(SensSeek ~ age0*groups + age0*DemPweight + (age0|PROC_CID), data = sample_dat)
```


# Time Varying Predictors

## Time Varying Predictors: Continuous
Next, we'll add in a time-varying predictor. Maybe it's not that our participants sensation seeking is moderated by early life experiences of jail or court-ordered community service. Instead, their sensation seeking is moderated by depression.  
How does this look?
\begin{itemize}
  \item \textbf{Level 1:} $Y_{ij} = \beta_{0j} + \beta_{1j}*time + \beta_{2j}*CESD + \varepsilon{ij}$
  \item \textbf{Level 2:} 
    \begin{itemize} 
      \item $\beta_{0j} = \gamma_{00} + U_{0j}$
      \item $\beta_{1j} = \gamma_{10} + U_{1j}$
      \item $\beta_{2j} = \gamma_{20}$
    \end{itemize}
\end{itemize}

###To Interaction or Not - That Is the Question
\begin{itemize}
  \item \textbf{Level 1:} $Y_{ij} = \beta_{0j} + \beta_{1j}*age0 + \beta_{2j}*CESD + \varepsilon{ij}$
  \item \textbf{Level 2:} 
    \begin{itemize} 
      \item $\beta_{0j} = \gamma_{00} + U_{0j}$
      \item $\beta_{1j} = \gamma_{10} + U_{1j}$
      \item $\beta_{2j} = \gamma_{20}$
    \end{itemize}
\end{itemize}
$$Y_{ij} =  \gamma_{00} + \gamma_{01} + U_{0j} + (\gamma_{10} + U_{1j})*age0 + \gamma_{20}*CESD$$

### Example: Does depression influence changes in sensation seeking over time?

\small
```{r, results = 'hide'}
modTV1 <- lmer(SensSeek ~ age0 + CESD + (age0|PROC_CID), data = sample_dat)
```

\small
```{r, echo = T}
summary(modTV1)
```

\normalsize

```{r, echo = T, eval = T}
# example for continuous
# note MEANS ARE AT AGE0 = 0
fixed.frame <- sample_dat %>%
  filter(age0 == 0) %>%
  summarise(mean = mean(CESD, na.rm = T), 
            sd = sd(CESD, na.rm = T))

fixed.frame <- 
  data.frame(
    expand.grid(
      # here, you add values for your time variable and predictors
      age0 = seq(0,10,2), 
      CESD = c(fixed.frame$mean-fixed.frame$sd,
                     fixed.frame$mean,
                     fixed.frame$mean+fixed.frame$sd))) %>%
  mutate(pred = predict(modTV1, newdata = ., re.form = NA))

fixed.frame %>%
  mutate(CESD = factor(CESD, levels = unique(CESD), labels = c("-1SD", "0SD", "1SD")),
         age = age0 + 14) %>%
  ggplot(aes(x = age, y = pred, color = CESD)) +
    geom_line(size = 2) + 
    labs(x = "Age", y = "Sensation Seeking Composite",
         title = "Continuous Time Varying Conditional Growth Models") +
    theme_classic() +
    theme(axis.text = element_text(face = "bold", size = rel(1.2)),
          axis.title = element_text(face = "bold", size = rel(1.2)),
          legend.title = element_text(face = "bold", size = rel(1.2)),
          plot.title = element_text(face = "bold", size = rel(1.2), hjust = .5))

```

## Time Varying Predictors: Categorical
Next, we'll add in a time-varying predictor. Maybe it's not that our participants sensation seeking is moderated by early life experiences of jail or court-ordered community service. Instead, their sensation seeking is moderated by depression.  
How does this look?
\begin{itemize}
  \item \textbf{Level 1:} $Y_{ij} = \beta_{0j} + \beta_{1j}*time + \beta_{2j}*depressed + \varepsilon{ij}$
  \item \textbf{Level 2:} 
    \begin{itemize} 
      \item $\beta_{0j} = \gamma_{00} + \gamma_{01} + U_{0j}$
      \item $\beta_{1j} = \gamma_{10} + U_{1j}$
      \item $\beta_{2j} = \gamma_{20}$
    \end{itemize}
\end{itemize}

\small
```{r}
# creating a dummy variable for time varying categorical depression
sample_dat <- sample_dat %>%
  mutate(depressed = 
           factor(ifelse(CESD <= 1.5, 0, 1), levels = c(0,1), 
                  labels = c("Depressed", "Not Depressed")))
modTV2 <- lmer(SensSeek ~ age0 + depressed + (age0|PROC_CID), 
               data = sample_dat)
summary(modTV2)
```

```{r}
fixed.frame <- 
  data.frame(expand.grid(age0 = seq(0,10,2), 
             depressed = c("Depressed","Not Depressed"))) %>%
  mutate(pred = predict(modTV2, newdata = ., re.form = NA))

fixed.frame %>%
  mutate(age = age0 + 14) %>%
  ggplot(aes(x = age, y = pred, color = depressed)) +
    geom_line(size = 2) + 
    lims(y = c(0,4)) +
    labs(x = "Age", y = "Sensation Seeking Composite",
         title = "2 Group Time Invariant Conditional Growth Models") +
    theme_classic() +
    theme(axis.text = element_text(face = "bold", size = rel(1.2)),
          axis.title = element_text(face = "bold", size = rel(1.2)),
          legend.title = element_text(face = "bold", size = rel(1.2)),
          plot.title = element_text(face = "bold", size = rel(1.2), hjust = .5))
```


\normalsize