---
title: "Conditional Models"
author: "Emorie D Beck"
date: "9/7/2017"
output: 
  pdf_document:
    keep_tex: yes
    includes:
            in_header:
                header.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, eval = T, cache = T, warning = F, message = F)
```

#Workspace
##Packages
```{r, results = 'hide', message = F, warning = F}
library(psych)
library(sjPlot)
library(broom)
library(lme4)
library(MuMIn)
library(merTools)
library(reghelper)
library(stargazer)
library(lsmeans)
library(multcompView)
library(plyr)
library(tidyverse)
```


# Background
## Basic Syntex
From last week:
\begin{itemize}
  \item \textbf{Level 1:} $Y_{ij} = \beta_{0j} + \varepsilon{ij}$
  \item \textbf{Level 2:} $\beta_{0j} = \gamma_{00} + U_{0j}$
\end{itemize}

## Sample Data
The National Longitudinal Study of Youths 1979 Child and Young Adult Sample (NLSYCYA) is a longitudinal study conducted by the National Bureau of Labor Statistics. The sample includes the children of the original 1979 sample. Here, we are going to use a subset of the more than 11,000 variables available that include the following.

Item Name   | Description                   | Time-Varying?
----------- | ----------------------------- | -------------
PROC_CID    | Participant ID                | No  
Dem_DOB     | Year of Date of Birth         | No
groups      | Jail, Community Service, None | No
DemPWeight  | Weight Percentile at age 10   | No
age         | Age of participant            | Yes
Year        | Year of Survey                | Yes
age0        | Age of participant (centered) | Yes
SensSeek    | Sensation-Seeking Composite   | Yes
CESD        | CESD Depression Composite     | Yes

```{r, eval = T, tidy = T}
data_path <- "https://github.com/longitudinal-data/1-descriptives-and-graphs-emoriebeck/raw/master/Conditional_Models"
load(url(paste(data_path, "sample.RData", sep = "/")))

head(sample_dat)
```


## Simple Growth Curve Model

\begin{itemize}
  \item \textbf{Level 1:} $Y_{ij} = \beta_{0j} + \beta_{1j}*time_{ij} + \varepsilon{ij}$
  \item \textbf{Level 2:} 
    \begin{itemize} 
      \item $\beta_{0j} = \gamma_{00} + U_{0j}$
      \item $\beta_{1j} = \gamma_{10} + U_{1j}$
    \end{itemize}
\end{itemize}

```{r, echo = F}
sample_dat %>%
  ggplot(aes(x = age, y = SensSeek)) + 
    geom_smooth(aes(group = PROC_CID), method = "lm", se = F, color = "gray", size = .2) +
    geom_smooth(method = "lm", se = F, color = "red", size = 2) +
    labs(x = "Age", y = "Sensation Seeking Composite", title = "Simple Growth Curve") +
    theme_classic() +
    theme(legend.position = "none",
          axis.text = element_text(face = "bold", size = rel(1.2)),
          axis.title = element_text(face = "bold", size = rel(1.2)),
          plot.title = element_text(face = "bold", size = rel(1.2), hjust = .5))
```

## In R
\small
```{r}
mod0 <- lmer(SensSeek ~ age0 + (1|PROC_CID), data = sample_dat)
```

\centering
\small
```{r, echo = F}
summary(mod0)
```

\normalsize
\raggedright

## Conditional Models: Adding Predictors
Let's see if we can better predict participants' change in sensation seeking over time by adding covariates.  

Predictor      | Continuous  | Categorical 
----------- | ----------- | -----------
Time Invariant  | Weight for Age  | Group 
Time Varying  | CESD Scores | Depression  


# Time Invariant Predictors
## Time Invariant Predictors: Continuous

The basic equation, specifying a random intercept and slope:    
\begin{itemize}
  \item \textbf{Level 1:} $Y_{ij} = \beta_{0j} + \beta_{1j}*time_{1j} + \varepsilon{ij}$
  \item \textbf{Level 2:} 
    \begin{itemize} 
      \item $\beta_{0j} = \gamma_{00} + \gamma_{01}*X_{2j} + U_{0j}$
      \item $\beta_{1j} = \gamma_{10} + \gamma_{11}*X_{2j} + U_{1j}$
    \end{itemize}
\end{itemize}

But we need to break this down to see that adding additional predictors results in interaction terms:  

$Y_{ij} = \gamma_{00} + \gamma_{01}*X_{2j} + U_{0j} + (\gamma_{10} + \gamma_{11}*X_{2j} + U_{1j})*X_{1j} + \varepsilon{ij}$
$Y_{ij} = \gamma_{00} + \gamma_{01}*X_{2j} + \gamma_{10}*X_{1j} + \textcolor{red}{\gamma_{11}*X_{2j}*X_{1j}} + U_{0j} + U_{1j}*X_{1j} + \varepsilon{ij}$

We can also fit this with intercepts depending on weight, but without the change (slope) dependent on weight:  
$Y_{ij} = \gamma_{00} + \gamma_{01}*X_{2j} + U_{0j} + (\gamma_{10} + U_{1j})*X_{1j} + \varepsilon{ij}$
$Y_{ij} = \gamma_{00} + \gamma_{01}*X_{2j} + \gamma_{10}*X_{1j} + U_{0j} + U_{1j}*X_{1j} + \varepsilon{ij}$

### Continuous Example - Weight for Age Percentile  

\small
```{r}
describe(sample_dat$DemPweight)
```

\normalsize

```{r, echo = F}
sample_dat %>%
  ggplot(aes(x = age, y = SensSeek)) + 
    geom_smooth(aes(group = PROC_CID, color = DemPweight), method = "lm", se = F, size = .2) +
    geom_smooth(method = "lm", se = F, color = "red", size = 1) +
    labs(x = "Age", y = "Sensation Seeking Composite", title = "Simple Growth Curve") +
    theme_classic() +
    theme(#legend.position = "none",
          axis.text = element_text(face = "bold", size = rel(1.2)),
          axis.title = element_text(face = "bold", size = rel(1.2)),
          plot.title = element_text(face = "bold", size = rel(1.2), hjust = .5))
```

\small
```{r,results='hide'}
# time invariant covariate with random intecept (with weight as covariate) 
# and slope (without weight as a covariate)
mod1a <- lmer(SensSeek ~ age0 + DemPweight + (age0|PROC_CID), 
              data = sample_dat)

summary(mod1a)

# time invariant predictor with random slope and intercept
mod1b <- lmer(SensSeek ~ age0 + DemPweight + age0*DemPweight + 
                (age0|PROC_CID), data = sample_dat)

summary(mod1b)
```

\normalsize

##Time Invariant Predictors: Categorical 
### Categorical Example - 2 level group

Let's start with the basic syntax:  

\begin{itemize}
  \item \textbf{Level 1:} $Y_{ij} = \beta_{0j} + \beta_{1j}*time_{1j} + \varepsilon{ij}$
  \item \textbf{Level 2:} 
    \begin{itemize} 
      \item $\beta_{0j} = \gamma_{00} + \gamma_{01}*X_{2j} + U_{0j}$
      \item $\beta_{1j} = \gamma_{10} + \gamma_{11}*X_{2j} + U_{1j}$
    \end{itemize}
\end{itemize}

Now let's swap that out for a 2 group sample from the present data:  

\begin{itemize}
  \item \textbf{Level 1:} $Y_{ij} = \beta_{0j} + \beta_{1j}*age0_{ij} + \varepsilon{ij}$
  \item \textbf{Level 2:} 
    \begin{itemize} 
      \item $\beta_{0j} = \gamma_{00} + \gamma_{01}*groupsNone + U_{0j}$
      \item $\beta_{1j} = \gamma_{10} + \gamma_{11}*groupsNone + U_{1j}$
    \end{itemize}
\end{itemize}

Variable  | D1  
--------- | ---
Jail      | 0 
None      | 1

And plot it.  
```{r, echo = F}
sample_dat %>% filter(groups != "CommServ") %>%
  ggplot(aes(x = age, y = SensSeek, color = groups)) +
    geom_smooth(aes(group = PROC_CID),method = "lm", se = F, color = "gray", size = .2) +
    geom_smooth(method = "lm", se = F, size = 1) +
    labs(x = "Age", y = "Sensation Seeking Composite",
         title = "2 Group Time Invariant Conditional Growth Models") +
    theme_classic() +
    theme(axis.text = element_text(face = "bold", size = rel(1.2)),
          axis.title = element_text(face = "bold", size = rel(1.2)),
          legend.title = element_text(face = "bold", size = rel(1.2)),
          plot.title = element_text(face = "bold", size = rel(1.2), hjust = .5))
```


And model it:  
\small
```{r}
mod2g <- lmer(SensSeek ~ age0 + groups + age0*groups + (age0|PROC_CID), 
              data = sample_dat %>% filter(groups != "CommServ"))
summary(mod2g)
```

\normalsize

### Categorical Example - 3 level group  

```{r, echo = F}
sample_dat %>%
  ggplot(aes(x = age, y = SensSeek, color = groups)) +
    geom_smooth(aes(group = PROC_CID), method = "lm", se = F, color = "gray", size = .2) +
    geom_smooth(method = "lm", se = F, size = 1) +
    labs(x = "Age", y = "Sensation Seeking Composite",
         title = "2 Group Time Invariant Conditional Growth Models") +
    theme_classic() +
    theme(axis.text = element_text(face = "bold", size = rel(1.2)),
          axis.title = element_text(face = "bold", size = rel(1.2)),
          legend.title = element_text(face = "bold", size = rel(1.2)),
          plot.title = element_text(face = "bold", size = rel(1.2), hjust = .5))
```


\begin{itemize}
  \item \textbf{Level 1:} $Y_{ij} = \beta_{0j} + \beta_{1j}*age0_{ij} + \varepsilon{ij}$
  \item \textbf{Level 2:} 
    \begin{itemize} 
      \item $\beta_{0j} = \gamma_{00} + \gamma_{01}*D1 + \gamma_{02}*D2 + U_{0j}$
      \item $\beta_{1j} = \gamma_{10} + \gamma_{11}*D1 + \gamma_{12}*D2 + U_{1j}$
    \end{itemize}
\end{itemize}

Variable  | D1  | D2  
--------- | --- | ---
Jail      | 0   | 0
None      | 1   | 0
CommServ  | 0   | 1  


\small
```{r}
mod3g <- lmer(SensSeek ~ age0 + groups + age0*groups + 
                (age0|PROC_CID), data = sample_dat)
summary(mod3g)
```


\normalsize

# Side Notes: Practical Applications
## Side Note: `lme4` helper functions  

```{r, eval = F}
vcov(mod2g)
VarCorr(mod2g)
fixef(mod2g)
head(ranef(mod2g)[[1]])
head(coef(mod2g)[[1]])
confint.merMod(mod2g, method = "boot")
reghelper::ICC(mod2g)
MuMIn::r.squaredGLMM(mod2g)
```

\small
```{r}
vcov(mod2g)
```


\small
```{r}
VarCorr(mod2g)
```


\small
```{r}
fixef(mod2g)
```


\small
```{r}
head(ranef(mod2g)[[1]])
```

```{r}
head(coef(mod2g)[[1]])
```

\small
```{r, warning = F, message = F}
confint.merMod(mod2g, method = "boot", nsim = 10)
```

All units of the random effects are in standard deviation units (which means you need to square them to get the variance!!)  
\begin{itemize}
  \item .sig01 = sd of random intercept = $\sqrt{\tau_{00}}$  
  \item .sig02 = correlation between slope and intercept = $\sqrt{\tau_{10}}$  
  \item .sig03 = sd of random slope = $\sqrt{\tau_{11}}$  
  \item .sigma = residual variance = $\hat{\sigma}$  
\end{itemize}


\small
```{r}
reghelper::ICC(mod2g)
```

***
\small
\textbf{Conditional $R^2$:} How much variance fixed + random effects explain  
\textbf{Marginal $R^2$:} how much variance the fixed effects explain  

[explained here](https://jonlefcheck.net/2013/03/13/r2-for-linear-mixed-effects-models/)  

```{r}
MuMIn::r.squaredGLMM(mod2g)
```

\normalsize

## Side Note: Creating MLM Tables
There are lots of helpful packages for this, including `stargazer` and `sjPlot`, which are demonstrated below.  
\small
```{r, eval = F}
stargazer::stargazer(mod2g)
sjPlot::sjt.lmer(mod2g)
```

\normalsize

The problem is that `stargazer()` doesn't include all the terms we want, and `sjt.lmer()` only renders html. Embedded in the `.Rmd` version of these slides is some code that should help you to extract the terms you need and create a table using `dplyr` and `tidyr` that you can render in \LaTeX using `stargazer`.

But let's understand where those variables came from. To do so, we'll use the `broom` package in R to grab the terms we need.  

Description                 | Math Notation 
--------------------------- | ------------------
Fixed Effect Intercept      | $\gamma_{00}$  
Fixed Effect Group Intercept  | $\gamma_{01}$  
Fixed Effect Age Slope      | $\gamma_{10}$  
Fixed Effect Group Slope    | $\gamma_{11}$  
Individual Random Intercepts  | $U_{0j}$  
Variance of Random Intercepts | $\tau_{00}$  
Random Age Slopes | $U_{10}$  
Variance of Random Age Slopes | $\tau_{11}$  
Correlation b/w Random Slopes and Intercepts  | $\tau_{10}$  
Residual Variance | $\hat{\sigma}^2$   
Intraclass Correlation  | ICC   
Conditional $R^2$           | $R^2_c$  
Marginal $R^2$              | $R^2_m$  


```{r, results='hide'}
broom::tidy(mod2g)
broom::glance(mod2g)
```

\small
```{r, echo = F}
broom::tidy(mod2g)
broom::glance(mod2g)
```

Below is code that *should* work for all models. Just run the function and save it as an `R` object. You can use this with `papaja` and the `apa_table()` function pretty easily. The trick is that if you are not using the papaja template, the proper LaTeX packages may not be loaded. You can get around this by attaching a .tex file calling the packages under "in_header: header.tex" in your YAML header. The YAML header of this .Rmd file contains the necessary syntax and the header.tex file with the proper packages.  
```{r, echo = T, results = 'hide'}
## here's some code to make a table. You shouldn't need to modify anything here 
# unless you add additional random effects terms
## fixed effects first ##
table_fun <- function(model){
    fixed <- broom::tidy(mod2g) %>% filter(group == "fixed") %>%
    select(term, estimate) 
  ## add random effects ##
  rand <- broom::tidy(mod2g) %>% filter(group != "fixed") %>%
    select(term, estimate)
  ## get confidence intervals ##
  CI <- data.frame(confint.merMod(mod2g, method = "boot", nsim = 10)) %>%
    mutate(term = rownames(.)) %>% setNames(c("lower", "upper", "term"))
  
  ## Get ICC & R2 values ##
  ICC <- reghelper::ICC(mod2g)
  R2 <- MuMIn::r.squaredGLMM(mod2g)
  
  ## format the fixed effects
  fixed <- fixed %>% left_join(CI %>% filter(!grepl(".sig", term))) %>%
    mutate(type = "Fixed Parts")
  
  rand <- rand %>%
    mutate(term = mapvalues(term, unique(term), 
            c("$\\tau{00}$", "$\\tau_{11}$", "$\\tau_{10}$", "$\\hat{\\sigma^2}$")),
           estimate = estimate^2) %>%
    left_join(
      CI %>% filter(grepl(".sig", term)) %>%
        mutate(term = mapvalues(term, unique(term), 
            c("$\\tau{00}$", "$\\tau_{10}$", "$\\tau_{11}$", "$\\hat{\\sigma^2}$")),
            lower = lower^2, upper = upper^2)) %>%
    mutate(type = "Random Parts")
  
  mod_terms <- tribble(
    ~term, ~estimate, ~type,
    "ICC", ICC, "Model Terms",
    "$R^2_m$", R2[1], "Model Terms",
    "$R^2_c$", R2[2], "Model Terms"
  )
  
  tab <- fixed %>%
    full_join(rand) %>%
    mutate(CI = sprintf("(%.2f, %.2f)", lower, upper)) %>%
    select(-lower, -upper) %>%
    full_join(mod_terms) %>%
    mutate(estimate = sprintf("%.2f", estimate)) %>%
    select(type, everything())
}
# you can use this with papaja and the apa_table function pretty easily
# the trick is that if you are not using the papaja template, the proper
# LaTeX packages may not be loaded. You can get around this by attaching
# a .tex file calling the packages under "in_header: header.tex" in your YAML
# header the YAML header of this .Rmd file contains the necessary syntax and 
# the header.tex file with the proper packages

tab <- table_fun(mod2g)
```

###Basic: `kable()`  
\small
```{r}
options(knitr.kable.NA = '')
knitr::kable(tab, caption = "Ugly MLM Table Example")
```


### More Advanced: `kable()` + `kableExtra`  
\small
```{r, results = 'asis'}
library(kableExtra)
options(knitr.kable.NA = '')
knitr::kable(tab %>% select(-type) %>%
    mutate(term = gsub("[()]", "", term)),
             caption = "Ugly MLM Table Example", format = "latex", 
             longtable = T, booktabs = T, escape = F) %>%
  group_rows("Fixed", 1,4) %>% 
  group_rows("Random", 5,9) %>%
  group_rows("Model", 9,11) %>%
  #kable_styling(latex_options = c("striped","repeat_header"),full_width = F)
  add_header_above(c(" ", "Model 1" = 2))
  
```


### Alternative: `papaja` + `apa_table()`  
\small
```{r, results = 'asis'}
papaja::apa_table(tab %>% select(-type),caption = "MLM Table Example", 
    na_string = "", stub_indents = list(Fixed = c(1:4), Random = c(5:11)))
```
\normalsize

##Side Note: Plotting
### Lazy Method: `sjPlot` + `sjt.int()`  

####Categorical
\small
```{r}
sjp.int(mod2g, type = "eff", p.kr = F, swap.pred = T)
```

####Continuous
\small
```{r}
sjp.int(mod1b, type = "eff", p.kr = F, swap.pred = T, mdrt.values = "meansd")
```

###More advanced: `expand.grid()` + `fixef()` + `ggplot2()`  
####Categorical
```{r, echo=T, eval=T}
# example for categorical
fixed.frame <- 
  data.frame(
    expand.grid(
      # here, you add values for your time variable and predictors
      age0 = seq(0,10,2), 
      groupsNone = c(0,1))) %>%
  # now take care of interactions and add an intercept
  mutate(`age0:groupsNone` = age0*groupsNone,
         Intercept = 1) %>%
  # reordering everything
  select(Intercept, everything())

# multiplying to get values for model frame
fixed.frame$value <- as.vector(as.matrix(fixed.frame) %*% fixef(mod2g))

fixed.frame %>%
  mutate(groups = factor(groupsNone, levels = c(0,1), labels = c("Jail", "None")),
         age = age0 + 14) %>%
  ggplot(aes(x = age, y = value, color = groups)) +
    geom_line(size = 2) + 
    labs(x = "Age", y = "Sensation Seeking Composite",
         title = "2 Group Time Invariant Conditional Growth Models") +
    theme_classic() +
    theme(axis.text = element_text(face = "bold", size = rel(1.2)),
          axis.title = element_text(face = "bold", size = rel(1.2)),
          legend.title = element_text(face = "bold", size = rel(1.2)),
          plot.title = element_text(face = "bold", size = rel(1.2), hjust = .5))

```

####Continuous

```{r, echo = T, eval = T}
# example for continuous
fixed.frame <- sample_dat %>%
  summarise(mean = mean(DemPweight, na.rm = T), 
            sd = sd(DemPweight, na.rm = T))

fixed.frame <- 
  data.frame(
    expand.grid(
      # here, you add values for your time variable and predictors
      age0 = seq(0,10,2), 
      DemPweight = c(fixed.frame$mean-fixed.frame$sd,
                     fixed.frame$mean,
                     fixed.frame$mean+fixed.frame$sd))) %>%
  # now take care of interactions and add an intercept
  mutate(`age0:DemPweight` = age0*DemPweight,
         Intercept = 1) %>%
  # reordering everything
  select(Intercept, everything())

# multiplying to get values for model frame
fixed.frame$value <- as.vector(as.matrix(fixed.frame) %*% fixef(mod1b))

fixed.frame %>%
  mutate(Weight = factor(DemPweight, levels = unique(DemPweight), labels = c("-1SD", "0SD", "1SD")),
         age = age0 + 14) %>%
  ggplot(aes(x = age, y = value, color = Weight)) +
    geom_line(size = 2) + 
    labs(x = "Age", y = "Sensation Seeking Composite",
         title = "Continuous Invariant Conditional Growth Models") +
    theme_classic() +
    theme(axis.text = element_text(face = "bold", size = rel(1.2)),
          axis.title = element_text(face = "bold", size = rel(1.2)),
          legend.title = element_text(face = "bold", size = rel(1.2)),
          plot.title = element_text(face = "bold", size = rel(1.2), hjust = .5))

```


\normalsize  

##Side Note: Comparisons with `lsmeans`
The `lsmeans` package has a lot of useful functions. They are listed below. Then I'll demonstrate them in turn.  
\small
```{r, eval = F}
# create a reference grid
ref.grid2g <- ref.grid(mod2g)
# create the lsmeans object
lsgroups   <- lsmeans(ref.grid2g, "groups")
# compact letter display
cld(lsgroups, alpha = .10)
# plot
plot(lsgroups)
# contrasts of the ref.grid object
contrast(ref.grid2g, method = "eff")
# comparisons
groups.sum <- summary(lsgroups, infer = c(TRUE,TRUE), 
                      level = .90, adjust = "bon", by = "groups")
```


```{r}
# create a reference grid
(ref.grid2g <- ref.grid(mod2g))
```


```{r}
# create the lsmeans object
(lsgroups   <- lsmeans(ref.grid2g, "groups"))
```


```{r}
# compact letter display
cld(lsgroups, alpha = .10)
```



```{r}
# plot
plot(lsgroups)
```



```{r}
# contrasts of the ref.grid object
contrast(ref.grid2g, method = "eff")
```



```{r}
# comparisons
(groups.sum <- summary(lsgroups, infer = c(TRUE,TRUE), 
          level = .90, adjust = "bon", by = "groups"))
```

# Time Varying Predictors

## Time Varying Predictors: Continuous
Next, we'll add in a time-varying predictor. Maybe it's not that our participants sensation seeking is moderated by early life experiences of jail or court-ordered community service. Instead, their sensation seeking is moderated by depression.  
How does this look?
\begin{itemize}
  \item \textbf{Level 1:} $Y_{ij} = \beta_{0j} + \beta_{1j}*time + \beta_{2j}*CESD + \varepsilon{ij}$
  \item \textbf{Level 2:} 
    \begin{itemize} 
      \item $\beta_{0j} = \gamma_{00} + \gamma_{01} + U_{0j}$
      \item $\beta_{1j} = \gamma_{10} + U_{1j}$
      \item $\beta_{2j} = \gamma_{20}$
    \end{itemize}
\end{itemize}

###To Interaction or Not - That Is the Question
\begin{itemize}
  \item \textbf{Level 1:} $Y_{ij} = \beta_{0j} + \beta_{1j}*age0 + \beta_{2j}*CESD + \varepsilon{ij}$
  \item \textbf{Level 2:} 
    \begin{itemize} 
      \item $\beta_{0j} = \gamma_{00} + \gamma_{01} + U_{0j}$
      \item $\beta_{1j} = \gamma_{10} + U_{1j}$
      \item $\beta_{2j} = \gamma_{20}$
    \end{itemize}
\end{itemize}
$$Y_{ij} =  \gamma_{00} + \gamma_{01} + U_{0j} + (\gamma_{10} + U_{1j})*age0 + \gamma_{20}*CESD$$

### Example: Does depression influence changes in sensation seeking over time?

\small
```{r, results = 'hide'}
modTV1 <- lmer(SensSeek ~ age0 + CESD + (age0|PROC_CID), data = sample_dat)
```

\small
```{r, echo = T}
summary(modTV1)
```

\normalsize

```{r, echo = T, eval = T}
# example for continuous
# note MEANS ARE AT AGE0 = 0
fixed.frame <- sample_dat %>%
  filter(age0 == 0) %>%
  summarise(mean = mean(CESD, na.rm = T), 
            sd = sd(CESD, na.rm = T))

fixed.frame <- 
  data.frame(
    expand.grid(
      # here, you add values for your time variable and predictors
      age0 = seq(0,10,2), 
      CESD = c(fixed.frame$mean-fixed.frame$sd,
                     fixed.frame$mean,
                     fixed.frame$mean+fixed.frame$sd))) %>%
  # now take care of interactions and add an intercept
  mutate(Intercept = 1) %>%
  # reordering everything
  select(Intercept, everything())

# multiplying to get values for model frame
fixed.frame$value <- as.matrix(fixed.frame) %*% as.vector(fixef(modTV1))

fixed.frame %>%
  mutate(CESD = factor(CESD, levels = unique(CESD), labels = c("-1SD", "0SD", "1SD")),
         age = age0 + 14) %>%
  ggplot(aes(x = age, y = value, color = CESD)) +
    geom_line(size = 2) + 
    labs(x = "Age", y = "Sensation Seeking Composite",
         title = "Continuous Time Varying Conditional Growth Models") +
    theme_classic() +
    theme(axis.text = element_text(face = "bold", size = rel(1.2)),
          axis.title = element_text(face = "bold", size = rel(1.2)),
          legend.title = element_text(face = "bold", size = rel(1.2)),
          plot.title = element_text(face = "bold", size = rel(1.2), hjust = .5))

```

## Time Varying Predictors: Categorical
Next, we'll add in a time-varying predictor. Maybe it's not that our participants sensation seeking is moderated by early life experiences of jail or court-ordered community service. Instead, their sensation seeking is moderated by depression.  
How does this look?
\begin{itemize}
  \item \textbf{Level 1:} $Y_{ij} = \beta_{0j} + \beta_{1j}*time + \beta_{2j}*depressed + \varepsilon{ij}$
  \item \textbf{Level 2:} 
    \begin{itemize} 
      \item $\beta_{0j} = \gamma_{00} + \gamma_{01} + U_{0j}$
      \item $\beta_{1j} = \gamma_{10} + U_{1j}$
      \item $\beta_{2j} = \gamma_{20}$
    \end{itemize}
\end{itemize}

\small
```{r}
# creating a dummy variable for time varying categorical depression
sample_dat <- sample_dat %>%
  mutate(depressed = 
           factor(ifelse(CESD <= 1.5, 0, 1), levels = c(0,1), 
                  labels = c("Depressed", "Not Depressed")))
modTV2 <- lmer(SensSeek ~ age0 + depressed + (age0|PROC_CID), 
               data = sample_dat)
summary(modTV2)
```


\normalsize
