---
title: "Assignment 5"
author: "Emorie D Beck"
date: "11/2/2017"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: cerulean
    highlight: tango
    code_folding: show
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F, cache = T)
```

#Workspace
##Packages
\footnotesize
```{r}
library(survey)
library(mi)
library(mitools)
library(psych)
library(MatchIt)
library(lavaan)
library(semTools)
library(lme4)
library(blavaan)
library(MuMIn)
library(gridExtra)
library(knitr)
library(kableExtra)
library(stargazer)
library(plyr)
library(stringr)
library(haven)
library(tidyverse)

data_path <- "~/Box Sync/network/other projects/PCLE Replication"
load(sprintf("%s/results/data.RData", data_path))
```

##Define Functions
```{r}
table_fun <- function(mod, target.var){
  df <- parameterestimates(mod) %>% 
      filter(lhs %in% target.var & op %in% c("~", "~1", "~~")) %>%
      unite(term, lhs, op, rhs, sep = "") %>%
      select(term, est, ci.lower, ci.upper) %>%
      mutate(type = ifelse(grepl("~~", term) == T, "random", "fixed")) %>%
      mutate(CI = sprintf("(%.2f, %.2f)", ci.lower, ci.upper),
             est = sprintf("%.2f", est)) %>%
     select(-ci.lower, -ci.upper) %>%
     gather(key = est, value = value, est, CI)
  
  
  fitmeas <- data.frame(fitmeasures(mod)) %>% 
    mutate(term = rownames(.), type = "fitmeas") %>% 
    dplyr::rename(value = fitmeasures.mod.) %>%
    mutate(value = sprintf("%.2f", value),
           est = "est") %>%
    filter(term %in% c("cfi", "rmsea", "chisq", "df"))
  
  df <- df %>% full_join(fitmeas)
}

mgm_table_fun <- function(mod, target.var){
  df <- parameterestimates(mod) %>% 
      filter(lhs %in% target.var & op %in% c("~", "~1", "~~")) %>%
      unite(term, lhs, op, rhs, sep = "") %>%
      select(group, term, est, ci.lower, ci.upper) %>%
      mutate(type = ifelse(grepl("~~", term) == T, "random", "fixed")) %>%
      mutate(CI = sprintf("(%.2f, %.2f)", ci.lower, ci.upper),
             est = sprintf("%.2f", est)) %>%
     select(-ci.lower, -ci.upper) %>%
     gather(key = est, value = value, est, CI)
  
  
  fitmeas <- data.frame(fitmeasures(mod)) %>% 
    mutate(term = rownames(.), type = "fitmeas", group = 1) %>% 
    dplyr::rename(value = fitmeasures.mod.) %>%
    mutate(value = sprintf("%.2f", value)) %>%
    filter(term %in% c("cfi", "rmsea", "chisq", "df"))
  
  df <- df %>% full_join(fitmeas)
}

mlm_table_fun <- function(model){
  fixed <- broom::tidy(model) %>% filter(group == "fixed") %>%
    select(term, estimate) 
  ## add random effects ##
  rand <- broom::tidy(model) %>% filter(group != "fixed") %>%
    select(term, estimate)
  ## get confidence intervals ##
  CI <- data.frame(confint.merMod(model, method = "boot", nsim = 10)) %>%
    mutate(term = rownames(.)) %>% setNames(c("lower", "upper", "term"))
  
  ## Get ICC & R2 values ##
  ICC <- reghelper::ICC(model)
  R2 <- MuMIn::r.squaredGLMM(model)
  
  ## format the fixed effects
  fixed <- fixed %>% left_join(CI %>% filter(!grepl(".sig", term))) %>%
    mutate(type = "Fixed Parts")
  
  rand <- rand %>%
    mutate(estimate = ifelse(grepl("cor", term) == T, estimate, estimate^2),
          term = mapvalues(term, unique(term), 
           c("$\\tau_{00}$", "$\\tau_{11}$", "$\\tau_{10}$", "$\\hat{\\sigma^2}$"))) %>%
    left_join(
      CI %>% filter(grepl(".sig", term)) %>%
        mutate(term = mapvalues(term, unique(term), 
            c("$\\tau_{00}$", "$\\tau_{10}$", "$\\tau_{11}$", "$\\hat{\\sigma^2}$")),
            lower = lower^2, upper = upper^2)) %>%
    mutate(type = "Random Parts")
  
  mod_terms <- tribble(
    ~term, ~estimate, ~type,
    "ICC", ICC, "Model Terms",
    "$R^2_m$", R2[1], "Model Terms",
    "$R^2_c$", R2[2], "Model Terms"
  )
  
  tab <- fixed %>%
    full_join(rand) %>%
    mutate(CI = sprintf("(%.2f, %.2f)", lower, upper)) %>%
    select(-lower, -upper) %>%
    full_join(mod_terms) %>%
    mutate(estimate = sprintf("%.2f", estimate)) %>%
    dplyr::rename(b = estimate) %>%
    select(type, everything())
  return(tab)
}
```


#Question 1:
Fit a measurement model to your constructs at one time point. Try out the different types of scaling discussed in class. What changes what stays the same?

```{r, results='asis'}
T1.mod <- '
  E.T1 =~ E.T1_1 + E.T1_2 + E.T1_3
  A.T1 =~ A.T1_1 + A.T1_2 + A.T1_3
  C.T1 =~ C.T1_1 + C.T1_2 + C.T1_3
  N.T1 =~ N.T1_1 + N.T1_2 + N.T1_3
  O.T1 =~ O.T1_1 + O.T1_2 + O.T1_3
'

T1.ec.mod <- '
  E.T1 =~ NA*E.T1_1 + e1*E.T1_1 + e2*E.T1_2 + e3*E.T1_3
  A.T1 =~ NA*A.T1_1 + a1*A.T1_1 + a2*A.T1_2 + a3*A.T1_3
  C.T1 =~ NA*C.T1_1 + c1*C.T1_1 + c2*C.T1_2 + c3*C.T1_3
  N.T1 =~ NA*N.T1_1 + n1*N.T1_1 + n2*N.T1_2 + n3*N.T1_3
  O.T1 =~ NA*O.T1_1 + o1*O.T1_1 + o2*O.T1_2 + o3*O.T1_3

  #model constraints for effect coding
  ## loadings must average to 1
  e1 == 3 - e2 - e3
  a1 == 3 - a2 - a3
  c1 == 3 - c2 - c3
  n1 == 3 - n2 - n3
  o1 == 3 - o2 - o3
'

mi_fun <- function(df, scale){
  if(scale == "marker"){
    cfa(T1.mod, data=df, missing ="FIML")
  } else if(scale == "fixed"){
    cfa(T1.mod, data=df, missing ="FIML", std.lv=TRUE)
  } else{
    cfa(T1.ec.mod, data=df, missing ="FIML")
  }
}

coefs <- paste(c("E","A","C","N","O"), rep("T1",5), sep = ".")

mm_dat <- bfi_wide %>% 
  gather(key = item, value = value, T1_1:T3_3) %>%
  unite(item, Trait, item, sep = ".") %>%
  spread(key = item, value = value) %>%
  nest() %>%
  mutate(mi.t1.marker = map(data, ~mi_fun(., "marker")),
         mi.t1.fixed  = map(data, ~mi_fun(., "fixed")),
         mi.t1.effect = map(data, ~mi_fun(., "effect")),
         mi.t1.marker.tab = map(mi.t1.marker, ~table_fun(., coefs)),
         mi.t1.fixed.tab = map(mi.t1.marker, ~table_fun(., coefs)),
         mi.t1.effect.tab = map(mi.t1.marker, ~table_fun(., coefs)))

mm_dat %>% unnest(mi.t1.marker.tab) %>% mutate(coding = "marker") %>%
  full_join(mm_dat %>% unnest(mi.t1.fixed.tab) %>% mutate(coding = "fixed")) %>%
  full_join(mm_dat %>% unnest(mi.t1.effect.tab) %>% mutate(coding = "effect")) %>%
  mutate(est = ifelse(type == "fitmeas", "est", est)) %>%
  filter(type %in% c("fixed", "fitmeas")) %>%
  unite(est, coding, est) %>%
  spread(key = est, value = value) %>%
  arrange(desc(type), term) %>%
  select(-type) %>%
  kable(., "html", booktabs = T, escape = F, digits = 2,
        caption = "Q1: CFA") %>%
  kable_styling(bootstrap_options = c("striped"),full_width = F) %>%
  group_rows("Latent Factors", 1,5) %>%
  group_rows("Model", 6,9)
```

# Question 2
What do the fit statistics say about your latent variable? Good/bad? Is your latent variable Just identified/saturdated, under identified or over identified?  

From this we see that our model is not great (CFI < .84, RMSEA > .07), but from our degrees of freedom, we see that our model is not overidentified.  
```{r, results = 'asis'}
fitmeas_fun <- function(fit){
  data.frame(fitmeasures(fit)) %>% 
    mutate(term = rownames(.), type = "fitmeas") %>% 
    dplyr::rename(est = fitmeasures.fit.) %>%
    filter(term %in% c("cfi", "rmsea", "chisq", "df"))
}

mm_dat <- mm_dat %>%
  mutate(fitmeas.t1.marker = map(mi.t1.marker, fitmeas_fun),
         fitmeas.t1.fixed  = map(mi.t1.fixed,  fitmeas_fun),
         fitmeas.t1.effect = map(mi.t1.effect, fitmeas_fun))

options(knitr.kable.NA = '')
mm_dat %>%  unnest(fitmeas.t1.marker) %>% mutate(model = "marker") %>%
  full_join(mm_dat %>% unnest(fitmeas.t1.fixed) %>% mutate(model = "fixed")) %>%
  full_join(mm_dat %>% unnest(fitmeas.t1.effect) %>% mutate(model = "effect")) %>%
  spread(key = model, value = est) %>%
  kable(., "html", digits = 2, booktabs = T, escape = F,
        caption = "Q2: Fit Statistics") %>%
  kable_styling(bootstrap_options = c("striped"),full_width = F) 
  
```

#Question 3:
Fit a longitudinal CFA model where you a) first correlate your latent factors across time and then b) a second model that predicts later times by a prevous time (ie auto regressive; t1 -> t2 -> t3). What are your conclusions? How does one differ from the other?  

We see a relatively large (> .75 ) autoregressive relationship between latent personality measures across time points, suggesting that measures were related over time.  

```{r, results='asis'}
mi.mod <- '
  T1 =~ T1_1 + T1_2 + T1_3 
  T2 =~ T2_1 + T2_2 + T2_3 
  T3 =~ T3_1 + T3_2 + T3_3 
'

cor.mod <- '
  T1 ~~ T2
  T2 ~~ T3
  T1 ~~ T3
'

ar.mod <- '
  T2 ~ T1
  T3 ~ T2
'

mi.cor.mod <- paste(mi.mod, cor.mod, sep = "\n\n")
mi.ar.mod  <- paste(mi.mod, ar.mod,  sep = "\n\n")

comp_fun <- function(df){
  df %>% gather(key = item, value = value, T1_1:T3_3) %>%
    separate(item, c("wave", "item"), sep = "_") %>%
    group_by(wave, PROC_SID) %>%
    summarize(value = mean(value, na.rm = T)) %>% 
    spread(key = wave, value = value)
}

mm.long_fun <- function(df, model){
  cfa(model, data = df, missing = "FIML")
}

mi_nested <- bfi_wide %>%
  group_by(Trait) %>%
  nest() %>%
  mutate(composites = map(data, comp_fun), 
         long.mi.cor = map(data, ~mm.long_fun(., mi.cor.mod)),
         long.mi.ar  = map(data, ~mm.long_fun(., mi.ar.mod)),
         long.mi.cor.tab = map(long.mi.cor, ~table_fun(., c("T1", "T2", "T3"))),
         long.mi.ar.tab = map(long.mi.ar, ~table_fun(., c("T1", "T2", "T3"))))


mi_nested %>% unnest(long.mi.cor.tab) %>% mutate(mod = "corr_lat") %>%
  full_join(mi_nested %>% unnest(long.mi.ar.tab) %>% mutate(mod = "autoreg")) %>%
  mutate(est = mapvalues(est, c("est", "CI"), c("b", "CI")),
         type = factor(type, levels = c("fixed", "random", "fitmeas"))) %>%
  unite(Trait, Trait, est, sep = ".") %>%
  spread(key = Trait, value = value) %>%
  arrange(type) %>%
  kable(., "html", booktabs = T, escape = F, digits = 2,
        caption = "Q4: 1st Order LGM") %>%
  kable_styling(bootstrap_options = c("striped"),full_width = F) %>%
  group_rows("Fixed", 1,8) %>%
  #group_rows("Intercept", 1,2) %>%
  group_rows("Random", 9, 18) %>%
  group_rows("Model", 19,25)
```

# Question 4
Fit a longitdinal growth model in SEM and in HLM. Compare and contrast the differences. 

The models are very similar. They both fit the data reasonably well and the fixed and random terms are within .05 each other. The only exception to this is the correlation between the random slopes and intercepts, which tended to be moderate and negative for mlm but very small and positive for SEM.  

```{r, results = 'asis'}

growth.mod <- '
  I =~ 1*T1 + 1*T2 + 1*T3
  S =~ 0*T1 + 1*T2 + 2*T3
'

growth_fun <- function(df, model){
  growth(model, data = df, missing = "FIML")
}

lmer_fun <- function(df){
  df <- df %>%
    gather(key = item, value = value, T1_1:T3_3, na.rm = T) %>%
    separate(item, c("wave", "item"), sep = "_") %>%
    mutate(wave = as.numeric(str_replace(wave, "T", ""))) %>%
    group_by(wave, PROC_SID) %>%
    summarize(composite = mean(value, na.rm = T),
              composite = ifelse(is.nan(composite) == T, NA_real_, composite)) %>%
    group_by(PROC_SID) %>% 
    mutate(n = n()) %>%
    filter(n > 1)
  lme4::lmer(composite ~ 1 + wave + (wave | PROC_SID), data = df)
}

mi_nested <- mi_nested %>% 
  mutate(sem.growth = map(composites, ~growth_fun(., growth.mod)),
         mlm.growth = map(data, lmer_fun),
         table.sem = map(sem.growth, ~table_fun(., c("I", "S"))),
         table.mlm = map(mlm.growth, mlm_table_fun))

mi_nested %>%
  unnest(table.mlm) %>%
  gather(key = param, value = value, b, CI) %>%
  unite(param, Trait, param, sep = ".") %>%
  spread(key = param, value = value) %>%
  mutate(type = factor(type, levels = c("Fixed Parts", "Random Parts", "Model Terms"))) %>%
  arrange(type) %>%
  kable(., "html", booktabs = T, escape = F, digits = 2,
        caption = "Q4: MLM") %>%
  kable_styling(bootstrap_options = c("striped"),full_width = F) %>%
  group_rows("Fixed", 1,2) %>%
  #group_rows("Intercept", 1,2) %>%
  group_rows("Random", 3,5) %>%
  group_rows("Model", 6,9)

mi_nested %>%
  unnest(table.sem) %>%
  mutate(est = mapvalues(est, c("est", "CI"), c("b", "CI")),
         type = factor(type, levels = c("fixed", "random", "fitmeas"))) %>%
  unite(Trait, Trait, est, sep = ".") %>%
  spread(key = Trait, value = value) %>%
  arrange(type) %>%
  kable(., "html", booktabs = T, escape = F, digits = 2,
        caption = "Q4: 1st Order LGM") %>%
  kable_styling(bootstrap_options = c("striped"),full_width = F) %>%
  group_rows("Fixed", 1,2) %>%
  #group_rows("Intercept", 1,2) %>%
  group_rows("Random", 3,5) %>%
  group_rows("Model", 6,9)
```


# Question 5
Constrain the residual variances to be equal. Does this change the fit of your model?  

The latent intercepts and slopes of the constrained model relative to the unconstrained model do not differ. THe random effects of the constrained model tend to be slightly smaller and have slighlty tighter confidence intervals. The CFI's of the constrained models are very slightly worse than the unsconstrained model, but the RMSEA is slightly better for the constrained v. unconstrained models.  
```{r, results = 'asis'}
resid.var <- '
  # equality of residuals 
  T1 ~~ u*T1
  T2 ~~ u*T2
  T3 ~~ u*T3
'

resid.growth.mod <- paste(growth.mod, resid.var, sep = "\n\n")

mi_nested <- mi_nested %>% 
  mutate(sem.growth.resid = map(composites, ~growth_fun(., resid.growth.mod)),
         table.resid = map(sem.growth.resid, ~table_fun(., c("I", "S"))))
         
mi_nested %>% unnest(table.resid) %>% mutate(res = "constrained") %>%
  full_join(mi_nested %>% unnest(table.sem) %>% mutate(res = "unconstrained")) %>%
  mutate(est = mapvalues(est, c("est", "CI"), c("b", "CI")),
         type = factor(type, levels = c("fixed", "random", "fitmeas"))) %>%
  unite(Trait, Trait, est, sep = ".") %>%
  spread(key = Trait, value = value) %>%
  arrange(type) %>%
  kable(., "html", booktabs = T, escape = F, digits = 2,
        caption = "Q5: 1st Order LGM with Constrained Residuals") %>%
  kable_styling(bootstrap_options = c("striped"),full_width = F) %>%
  group_rows("Fixed", 1,4) %>%
  #group_rows("Intercept", 1,2) %>%
  group_rows("Random", 5,10) %>%
  group_rows("Model", 11,18)
```

# Question 6
Contrain your slope to be fixed, not random. How does this change your model?  

The fixed parameters are almost completely identical and the random parameters are also very close, although the variance of the random effects is slightly larger when slopes are fixed. The biggest change comes in the fit statistics where we have fewer degrees of freedom with the fixed slope and slighlty worse fit indices. Fixing the slope does not seem to add much in this case.  

```{r, results = 'asis'}
fixed.slope <- '
  S~0*S
'
fs.growth.mod <- paste(resid.growth.mod, fixed.slope, sep = "\n\n")

mi_nested <- mi_nested %>% 
  mutate(sem.growth.fs = map(composites, ~growth_fun(., fs.growth.mod)),
         table.fs = map(sem.growth, ~table_fun(., c("I", "S"))))

mi_nested %>% unnest(table.fs) %>% mutate(res = "fixed_slope") %>%
  full_join(mi_nested %>% unnest(table.resid) %>% mutate(res = "constrained")) %>%
  mutate(est = mapvalues(est, c("est", "CI"), c("b", "CI")),
         type = factor(type, levels = c("fixed", "random", "fitmeas"))) %>%
  unite(Trait, Trait, est, sep = ".") %>%
  spread(key = Trait, value = value) %>%
  arrange(type) %>%
  kable(., "html", booktabs = T, escape = F, digits = 2,
        caption = "Q6: 1st Order LGM with Fixed Slopes") %>%
  kable_styling(bootstrap_options = c("striped"),full_width = F) %>%
  group_rows("Fixed", 1,4) %>%
  #group_rows("Intercept", 1,2) %>%
  group_rows("Random", 5,10) %>%
  group_rows("Model", 11, 18)
```

# Question 7
Change the time metric in your SEM growth model. How does that change your estimates? Does it change your fit statistics?  

The slopes of the new time metric are flatter than the old time metric because a 1 unit change now signals 1 year, which is .25 waves. Fit statistics are identical.    
```{r, results = 'asis'}
nt.growth.mod <- '
  I =~ 1*T1 + 1*T2 + 1*T3
  S =~ 0*T1 + 4*T2 + 8*T3
'

mi_nested <- mi_nested %>% 
  mutate(sem.growth.nt = map(composites, ~growth_fun(., nt.growth.mod)),
         table.nt = map(sem.growth.nt, ~table_fun(., c("I", "S"))))

mi_nested %>% unnest(table.nt) %>% mutate(time = "New Time") %>%
  full_join(mi_nested %>% unnest(table.sem) %>% mutate(time = "OG Time")) %>%
  mutate(est = mapvalues(est, c("est", "CI"), c("b", "CI")),
         type = factor(type, levels = c("fixed", "random", "fitmeas"))) %>%
  unite(Trait, Trait, est, sep = ".") %>%
  spread(key = Trait, value = value) %>%
  arrange(type) %>%
  kable(., "html", booktabs = T, escape = F, digits = 2,
        caption = "Q7: 1st Order LGM with Different Time Metric") %>%
  kable_styling(bootstrap_options = c("striped"),full_width = F) %>%
  group_rows("Fixed", 1,4) %>%
  #group_rows("Intercept", 1,2) %>%
  group_rows("Random", 5,10) %>%
  group_rows("Model", 11,18)
```

# Question 8
Try a different type of estimation (see lavaan tutorial for details). How does that change your model?  

None of the parameters, including fit indices, seem to change.  
```{r, results='asis'}
growth_fun_dwls <- function(df, model){
  growth(model, data = df, missing = "FIML", estimator = "DWLS")
}

mi_nested <- mi_nested %>% 
  mutate(sem.growth.resid.dwls = map(composites, ~growth_fun(., resid.growth.mod)),
         table.dwls = map(sem.growth.resid.dwls, ~table_fun(., c("I", "S"))))

mi_nested %>% unnest(table.dwls) %>% mutate(estim = "DWLS") %>%
  full_join(mi_nested %>% unnest(table.resid) %>% mutate(estim = "default")) %>%
  mutate(est = mapvalues(est, c("est", "CI"), c("b", "CI")),
         type = factor(type, levels = c("fixed", "random", "fitmeas"))) %>%
  unite(Trait, Trait, est, sep = ".") %>%
  spread(key = Trait, value = value) %>%
  arrange(type) %>%
  kable(., "html", booktabs = T, escape = F, digits = 2,
        caption = "Q8: 1st Order LGM with DWLS Estimator") %>%
  kable_styling(bootstrap_options = c("striped"),full_width = F) %>%
  group_rows("Fixed", 1,4) %>%
  #group_rows("Intercept", 1,2) %>%
  group_rows("Random", 5,10) %>%
  group_rows("Model", 11,18)
```

# Question 9
Provide semplots for each of the models  
```{r}

```


# Question 10
Test measurement invariance across time for your construct. Can you run growth models? If there is evidence of non-invariance, what seems to be the problem?  
```{r, results = 'hide'}
constrainedVar <- list(
  paste("T1",1:3, sep = "_"),
  paste("T2",1:3, sep = "_"),
  paste("T3",1:3, sep = "_")
)

mi.long_fun <- function(df, model){
  longInvariance(model, auto=1, constrainAuto=TRUE, 
                 varList=constrainedVar, data=df, missing ="FIML")
}

extract_fun <- function(fit){
  ldply(fit, fitmeasures) %>%
    select(.id, cfi, rmsea) %>% 
    mutate(cfi.delta = 
      ifelse(.id == "fit.loadings", cfi[.id == "fit.configural"] - cfi[.id == "fit.loadings"],
      ifelse(.id == "fit.thresholds", cfi[.id == "fit.loadings"] - cfi[.id == "fit.thresholds"],
      ifelse(.id == "fit.means", cfi[.id == "fit.thresholds"] - cfi[.id == "fit.means"], NA))),
      rmsea.delta = 
      ifelse(.id == "fit.loadings", cfi[.id == "fit.configural"] - cfi[.id == "fit.loadings"],
      ifelse(.id == "fit.thresholds", cfi[.id == "fit.loadings"] - cfi[.id == "fit.thresholds"],
      ifelse(.id == "fit.means", cfi[.id == "fit.thresholds"] - cfi[.id == "fit.means"], NA))))
}

mi_nested <- mi_nested %>% 
  mutate(long.mi = map(data, ~mi.long_fun(., mi.mod)),
         res = map(long.mi, extract_fun))
```

```{r, results='asis'}
mi_nested %>% unnest(res) %>%gather(key = res, value = value, cfi:rmsea.delta) %>%
  spread(key = res, value = value) %>%
  arrange(.id, Trait) %>%
  kable(., "html", booktabs = T, digits = 2,
        caption = "Q10: Longitudinal Measurment Invariance") %>%
  kable_styling(bootstrap_options = c("striped"),full_width = F) %>%
  collapse_rows()
```


# Question 11
Fit a second order growth model. Compare and contrast the estimates with the normal latent growth model.  

The intercepts in the first order growth curve models are larger than the 2nd order latent estimates. The variance terms are very close in both models, although they are slighlty larger in the 1st order than the second order model. Model fit indices are better for the 2nd order model.  

```{r, results='asis'}
base.growth.mod <- '
T1 =~ NA*T1_1 + (lambda1)*T1_1 + (lambda2)*T1_2 + (lambda3)*T1_3 
T2 =~ NA*T2_1 + (lambda1)*T2_1 + (lambda2)*T2_2 + (lambda3)*T2_3 
T3 =~ NA*T3_1 + (lambda1)*T3_1 + (lambda2)*T3_2 + (lambda3)*T3_3 


###intercepts
T1_1 ~ (nu1)*1
T1_2 ~ (nu2)*1
T1_3 ~ (nu3)*1
T2_1 ~ (nu1)*1
T2_2 ~ (nu2)*1
T2_3 ~ (nu3)*1
T3_1 ~ (nu1)*1
T3_2 ~ (nu2)*1
T3_3 ~ (nu3)*1

# T1_1 ~ 1
# T1_2 ~ 1
# T1_3 ~ 1
# T2_1 ~ 1
# T2_2 ~ 1
# T2_3 ~ 1
# T3_1 ~ 1
# T3_2 ~ 1
# T3_3 ~ 1

####variances/covariances
# item 1
T1_1 ~~ T2_1
T1_1 ~~ T3_1
T2_1 ~~ T3_1

# item 2
T1_2 ~~ T2_2
T1_2 ~~ T3_2
T2_2 ~~ T3_2

# item 3
T1_3 ~~ T2_3
T1_3 ~~ T3_3
T2_3 ~~ T3_3

# I ~~ I
# S ~~ S
# I ~~ S
# I ~ 1
# S ~ 1
# T1 ~ 0*1
# T2 ~ 0*1
# T3 ~ 0*1

#model constraints
lambda1 == 3 - lambda2 - lambda3 
(nu1) == 0 - (nu2) - (nu3)

'

def.latent <- '
  I =~ 1*T1 + 1*T2 + 1*T3
  S =~ 0*T1 + 1*T2 + 2*T3
'

growth.mod <- paste(base.growth.mod, def.latent, sep = "\n\n")

mi_nested <- mi_nested %>% 
  mutate(sem.sec.growth = map(data, ~growth_fun(., growth.mod)),
         table.sec = map(sem.sec.growth, ~table_fun(., c("I", "S"))))

mi_nested %>% unnest(table.sec) %>% mutate(order = "2nd") %>%
  full_join(mi_nested %>% unnest(table.resid) %>% mutate(order = "1st")) %>%
  mutate(est = mapvalues(est, c("est", "CI"), c("b", "CI")),
         type = factor(type, levels = c("fixed", "random", "fitmeas"))) %>%
  unite(Trait, Trait, est, sep = ".") %>%
  spread(key = Trait, value = value) %>%
  arrange(type) %>%
  kable(., "html", booktabs = T, escape = F, digits = 2,
        caption = "2nd Order LGM") %>%
  kable_styling(bootstrap_options = c("striped"),full_width = F) %>%
  group_rows("Fixed", 1,4) %>%
  #group_rows("Intercept", 1,2) %>%
  group_rows("Random", 5,10) %>%
  group_rows("Model", 11,18)
```


# Question 12
Fit a series of multiple group models. Constrain some parameters and compare the fit.

```{r, results='asis'}
mgm_fun <- function(df, model){
  growth(model, data = df, missing = "FIML", group = "le.group")
}

mgm.nested <- bfi_wide %>%
  full_join(le_dat %>% select(Event, PROC_SID, le.group)) %>%
  filter(!is.na(le.group) & Event == "Divorce") %>%
  group_by(Trait, Event) %>%
  nest() %>%
  mutate(mgm = map(data, ~mgm_fun(., growth.mod)),
         table.mgm = map(mgm, ~mgm_table_fun(., c("I", "S"))))

mgm.nested %>%
  unnest(table.mgm) %>%
  mutate(est = mapvalues(est, c("est", "CI"), c("b", "CI")),
         type = factor(type, levels = c("fixed", "random", "fitmeas"))) %>%
  unite(Trait, Trait, group, sep = ".") %>%
  spread(key = Trait, value = value) %>%
  arrange(type) %>%
  kable(., "html", booktabs = T, escape = F, digits = 2,
        caption = "Multiple Groups Models 2nd Order LGM") %>%
  kable_styling(bootstrap_options = c("striped"),full_width = F) %>%
  group_rows("Fixed", 1,4) %>%
  #group_rows("Intercept", 1,2) %>%
  group_rows("Random", 5,10) %>%
  group_rows("Model", 11,14)
```

